---
title: "Take-Home Exercise 01"
author: "Dew Stella Chan"
date: "2025-02-01"
date-modified: "last-modified"
execute:
  echo: true
  eval: true
  warning: false
  freeze: true
---

This quota document shows the visual analytics conducted for the Take Home Exercise 01, The data set chosen is the [Heart Attack in Japan Youth Vs Adult](https://www.kaggle.com/code/radhikaarajdev/heart-attack-risk-youth-vs-adults-in-japan) from Kaggle.

## Required Task

The main task is to develop graphical visuals for a media company on an article for the possible factors leading to heart attack trends in Japanese Youth.

The plan is to carry out exploratory and confirmatory data analysis to confirm the observed trend before doing up the visuals required for the graphic.

The goal is to find out if there are any observable trends and health or lifestyle factors leading to the occurrence of heart attack in Japanese youth.

For the purpose of this analysis, I define Japanese youth as individuals under the age of 35, following the definition provided in the article on [Youth Employment Policies in Japan](https://www.jil.go.jp/english/jli/documents/2023/041-05.pdf). The data set is catogrised into 4 main groups as follows:

1)  Youths who have experienced a heart attack (interest group),
2)  Youths who have not experienced a heart attack,
3)  Adults who have experienced a heart attack,
4)  Adults who have not experienced a heart attack.

Population with age above 65, they will be omitted from this study as age above 65 and above are considered as elderly based on the [Wikipedia.](https://en.wikipedia.org/wiki/Aging_of_Japan#:~:text=Article,as%20South%20Korea%20and%20China.)

## 1.0 Loading of Packages

The following code chunk indicates the list of packages used for this Take Home Exercise 01.

```{r}
pacman::p_load(haven, SmartEDA, tidyverse, tidymodels, ggdist, ggridges, ggthemes,
               colorspace, gridExtra, ggstatsplot, GGally, 
               readxl, performance, parameters, car,aod, ggExtra, ggiraph, corrplot)
```

## 2.0 Importing of Data and Data Preparation.

### 2.1 Importing data.

As the data is in csv file format, the data is loaded using read_csv function.

```{r}
heart <- read_csv("data/Ex01/japan_heart_attack_dataset.csv")
```

### 2.2 Data Preparation

As the metadata did not specify the data in the extra column 1 to 15, hence we will not be able the data in the "Extra_column 1 to 15"

The first step of data preparation is to remove them first, and remove population who are above 64.

```{r}
heart <- heart %>% select (1:17) %>% filter (Age < 65)
```

The summary and head functions are used to view and examine the data before proceeding further.

```{r}
summary(heart)
```

```{r}
head (heart)
```

### 2.3 Check for missing values

is.na function is used to check if the data contain any missing values,if there are, they are being added.

```{r}
colSums(is.na(heart))
```

After removing the Extra columns, we continue by examining again the data using the summary and head functions to ensure that the extra columns are successfully removed.

```{r}
summary(heart)
```

```{r}
head(heart)
```

```{r}
heart <- heart %>%
  mutate(Age_Grp = ifelse(Age > 35, "Adult", "Youth"))
```

### 2.4 Transforming the data to the 4 main sub group.

It was observed that the data consist two types of variables: namely continuous and categorical.

As mentioned in section 1.0, the data will be categorised into the 4 main group.

Given that the continuous data are consist of variables such as heart rate, cholesterol level, Systolic and Diastolic Blood Pressure, they are being categorized to established category . For this purpose of the analysis, we will use the range indicate in the following sites and with the help of Microsoft Co-pilot.

1.  [Body Mass Index (BMI)](https://www.cdc.gov/bmi/adult-calculator/bmi-categories.html?form=MG0AV3)

2.  [Blood pressure (BP)](https://www.heart.org/en/health-topics/high-blood-pressure?form=MG0AV3)

3.  [Cholesterol Levels](https://www.heart.org/en/health-topics/cholesterol/about-cholesterol/what-your-cholesterol-levels-mean)

4.  [Heart Rate](https://www.mayoclinic.org/healthy-lifestyle/fitness/expert-answers/heart-rate/faq-20057979)

Using the following codes, the data are categorised into 4 main groups and the health indicators are categorized into the established data.

```{r}
HA <- heart %>%
  mutate(Event_Grp = ifelse(Age_Grp == "Adult" & Heart_Attack_Occurrence == "Yes", "Adult_HA",
                    ifelse(Age_Grp == "Adult" & Heart_Attack_Occurrence == "No", "Adult_No_HA",
                    ifelse(Age_Grp == "Youth" & Heart_Attack_Occurrence == "Yes", "Youth_HA",
                    ifelse(Age_Grp == "Youth" & Heart_Attack_Occurrence == "No", "Youth_No_HA",NA)))))

HA <- HA %>%
  mutate(BMI_cat = case_when( 
    BMI < 18.5 ~ "Underweight",
    BMI >= 18.5 & BMI < 25 ~ "Normal",
    BMI >= 25 & BMI < 30 ~ "Overweight",
    BMI >= 30 ~ "Obese",
    TRUE ~ NA_character_
  ))

HA <- HA %>%
  mutate(BP_cat = case_when(
    Systolic_BP < 120 & Diastolic_BP < 80 ~ "Normal",
    (Systolic_BP >= 120 & Systolic_BP < 130) & Diastolic_BP < 80 ~ "Elevated",
    (Systolic_BP >= 130 & Systolic_BP < 140) | (Diastolic_BP >= 80 & Diastolic_BP < 90) ~ "High Stage 1",
    Systolic_BP >= 140 | Diastolic_BP >= 90 ~ "High Stage 2",
    TRUE ~ NA_character_
  ))

HA <- HA %>%
  mutate(Cholesterol_Level_cat = case_when(
    Cholesterol_Level < 200 ~ "Desirable",
    Cholesterol_Level >= 200 & Cholesterol_Level < 240~ "Borderline High",
    Cholesterol_Level >= 240 ~ "High",
    TRUE ~ NA_character_
  ))

HA <- HA %>%
  mutate(Stress_Level_cat = case_when(
    Stress_Levels < 3 ~ "Low",
    Stress_Levels >= 3 & Stress_Levels < 6 ~ "Moderate",
    Stress_Levels >= 6 ~ "High",
    TRUE ~ NA_character_
  ))


HA <- HA %>% 
  mutate(Heart_Rate_cat = case_when(
    Heart_Rate < 70 ~ "Below Normal",
    Heart_Rate >=  70 & Heart_Rate <= 100 ~ "Normal",
    Heart_Rate  > 100 ~ "Elevated",
    TRUE ~ NA_character_
  ))

```

Examine of missing data after recoding

note that there are 277 recids of BMI data and 108 of chelesterol records not being

```{r}
missing_values <- colSums(is.na(HA))
print(missing_values)
```

```{r}
summary(HA)
```

### 2.5 Re-coding Catological data for Logistic Regression

```{r}
HA <- HA %>%
  mutate(Heart_Attack_Occurrence_num = case_when(
      Heart_Attack_Occurrence == "Yes" ~ 1,
      Heart_Attack_Occurrence == "No" ~ 0,
      TRUE ~ NA_real_),
    Gender_num = case_when(
      Gender == "Male" ~ 1,
      Gender == "Female" ~ 0,
      TRUE ~ NA_real_),
    Region_num = case_when(
      Region == "Urban" ~ 1,
      Region == "Rural" ~ 0,
      TRUE ~ NA_real_),
    Smoking_History_num = case_when(
      Smoking_History == "Yes" ~ 1,
      Smoking_History == "No" ~ 0,
      TRUE ~ NA_real_),
    Diabetes_History_num = case_when(
      Diabetes_History == "Yes" ~ 1,
      Diabetes_History == "No" ~ 0,
      TRUE ~ NA_real_),
    Hypertension_History_num = case_when(
      Hypertension_History == "Yes" ~ 1,
      Hypertension_History == "No" ~ 0,
      TRUE ~ NA_real_),
    Family_History_num = case_when(
      Family_History == "Yes" ~ 1,
      Family_History == "No" ~ 0,
      TRUE ~ NA_real_),
    Physical_Activity_num = case_when(
      Physical_Activity == "High" ~ 0,
      Physical_Activity == "Moderate" ~ 1,
      Physical_Activity == "Low" ~ 2,
      TRUE ~ NA_real_),
    Diet_Quality_num = case_when(
      Diet_Quality == "Good" ~ 0,
      Diet_Quality == "Average" ~ 1,
      Diet_Quality == "Poor" ~ 2,
      TRUE ~ NA_real_),
    Alcohol_Consumption_num = case_when(
      Alcohol_Consumption == "None" ~ 0,
      Alcohol_Consumption == "Low" ~ 1,
      Alcohol_Consumption == "Moderate" ~ 2,
      Alcohol_Consumption == "High" ~ 3,
      TRUE ~ NA_real_),
    Stress_Level_cat_num = case_when(
      Stress_Level_cat == "Low" ~ 0,
      Stress_Level_cat == "Moderate" ~ 1,
      Stress_Level_cat == "High" ~ 2,
      TRUE ~ NA_real_),
    BMI_cat_num = case_when(
      BMI_cat == "Underweight" ~ 0,
      BMI_cat == "Normal" ~ 1,
      BMI_cat == "Overweight" ~ 2,
      BMI_cat == "Obese" ~ 2,
      TRUE ~ NA_real_),
    Heart_Rate_cat_num = case_when(
      Heart_Rate_cat  == "Below Normal" ~ 0,
      Heart_Rate_cat  == "Normal" ~ 1,
      Heart_Rate_cat  == "Elevated" ~ 2,
      TRUE ~ NA_real_),
    BP_cat_num = case_when(
      BP_cat  == "Normal" ~ 0,
      BP_cat  == "Elevated" ~ 1,
      BP_cat  == "High Stage 1" ~ 2,
      BP_cat  == "High Stage 2" ~ 3,
      TRUE ~ NA_real_),
    Cholesterol_Level_cat_num = case_when(
      Cholesterol_Level_cat  == "Desirable" ~ 0,
      Cholesterol_Level_cat == "Borderline High" ~ 1,
      Cholesterol_Level_cat  == "High" ~ 2,
      TRUE ~ NA_real_))


```

Checking of data after re-coding.

```{r}
missing_values <- colSums(is.na(HA))
print(missing_values)
```

### 2.6 Prepare of varies data frame for easy of infographic.

```{r}
HA_Yes <- filter(HA, Heart_Attack_Occurrence == "Yes")
HA_adult <- filter(HA, Age_Grp == "Adult")
HA_youth <- filter(HA, Age_Grp == "Youth")
HA_youth_yes <- filter(HA_youth, Heart_Attack_Occurrence == "Yes")
```

## 3.0 Exploratory Data Analysis (EDA)

We start off by understanding the distribution of the variables and examine if there were any observable trends to be tested out using confirmatory analysis.

### 3.1 Visualising Distribution of the Categorical Variables

The aim of this task is to examine if there are any trends in categorical data in the occurrence of heart attack in Japanese population.

The code below is extract our the categorical data and summaries into the occurrence of heart attack by Heart occurrence Age Groups and Event Groups.

The data are processed using the following codes to allow flexibility when exploring the data set.Hence HA is being used instead of the subgroups. Those data sets would be used in the subsequent visuslisation.

```{r}
categorical_columns <- HA %>% select(where(~is.character(.x) || is.factor(.x)))

count_occurrences <- categorical_columns %>%
  pivot_longer(cols= - c(Heart_Attack_Occurrence, Age_Grp, Event_Grp), names_to = "Category", values_to = "Value") %>%
  group_by(Category, Value, Age_Grp, Heart_Attack_Occurrence, Event_Grp) %>%
  summarise(Occurrences = n(), .groups = 'drop')
```

```{r}
#| fig-width: 20 #to widen the space
#| fig-height: 16 #to lengthen the graph

bar_plots <- count_occurrences %>%
  ggplot(aes(x = Value, y = Occurrences, fill = Heart_Attack_Occurrence)) + 
  geom_bar(stat = "identity", position = position_dodge()) +  
  facet_wrap(~ Category, scales = "free") +
  theme_minimal() +
  labs(title = "Distribution of Categorical Variables of The Study Population.",
       x = "Unique Values",
       y = "Count") +
  scale_fill_manual(values = c("Yes" = "red", "No" = "#4682B4"))

bar_plots
```

::: Callout-note
The above chart shows the data distribution for the categorical variables of the data set.

It includes the entire data set of the risk factors with whom had have heart attack verse whom had not have attack in Japan in all age group.

Before visualizing the data, one would expect [risk factors](https://www.mayoclinic.org/diseases-conditions/heart-attack/symptoms-causes/syc-20373106) such as high cholesterol level, elevated heart rate and with smoking history, will likely result in more occurrence of heart attack in but solely by looking at the distribution of the data, this may not be true.

The main take away from the analysis is that it seems there is no **sole risk factor** which will lead to heart attack with the Japanese population sampled for this data set.
:::

#### 3.1.1 Closer Examination on the Distribution Categorical Variables in different Age group in Youth vs Adult

```{r}
#| fig-width: 20 #to widen the space
#| fig-height: 16 #to lengthen the graph
count_occurrences_yes <- filter(count_occurrences,  Heart_Attack_Occurrence == "Yes")

bar_plots_yes <- count_occurrences_yes %>%
  ggplot(aes(x = Value, y = Occurrences, fill = Age_Grp)) + 
  geom_bar(stat = "identity", position = position_dodge()) + 
  facet_wrap(~ Category, scales = "free") +
  theme_minimal() +
  labs(title = "Distribution of Categorical Variables of subjects with Heart Attack Across Age Groups",
       y = "Count") +
scale_fill_manual(values = c("Youth" = "#FF1493", "Adult" = "#FFA07A"))

bar_plots_yes
```

::: Callout-note
Due to the study population, contains more subjects with heart attack, it is good to visualise the distribution across the age groups with heart attack. This may provide some interesting insights.

The charts above show the distribution of the categorical variables for study subjects who had heart attack across different age groups.

At a glance, the general distribution across categorical variables may seem to be similar across both age groups but at a closer examination. the following observations could be seen:

-   BP_cat "high stage 1" in adults has higher occurrence in heart attack when compared to youth.

-   In youth, the occurrence of heart attack across gender while in adult a slightly in male verse female.

-   There seems to be have more occurrence in heart attack occurring in urban setting than rural setting.

-   It seem to have association between BMI and heart attack with more slight more occurrence in overweight subjects as compared to other BMI category.

-   It seems that there is association with moderate stress level with high proportion in both age groups.

From the distribution, it seems that there is no sole factor which stand out in the youth population which will result in heart attack, confirmatory analysis would be need to confirm the analysis or explanatory model using logistic regression could be carried out to further explain risk factors.
:::

#### 3.1.2 Closer Examination on the Distribution Categorical Variables in the Youth Age group.

```{r}
#| fig-width: 20 #to widen the space
#| fig-height: 16 #to lengthen the graph
count_occurrences_youth <- filter(count_occurrences,  Age_Grp == "Youth")

bar_plots_yes <- count_occurrences_youth %>%
  ggplot(aes(x = Value, y = Occurrences, fill = Heart_Attack_Occurrence)) + 
  geom_bar(stat = "identity", position = position_dodge()) + 
  facet_wrap(~ Category, scales = "free") +
  theme_minimal() +
  labs(title = "Distribution of Categorical Variables of subjects with Heart Attack in Youth",
       y = "Count") +
scale_fill_manual(values = c("Yes" = "#FF1493", "No" = "#87CEEB"))

bar_plots_yes
```

::: Callout-note
Similar to the bar charts in section 3.1, there is no *sole risk factor* which stands out in the categorical variables which lead to high occurrence in the youth population.
:::

### 3.2 Visualising Distribution of the Continuous Variables

The following codes is to prepared the data to plot histogram to visualise the continuous data.

The data is prepared this manner so that it will provide flexibility in showing data.

```{r}
contin_col_HA <- heart %>% select(where(is.numeric), Heart_Attack_Occurrence, Age_Grp)

contin_col_HA_youth <- filter(contin_col_HA, Age_Grp == "Youth")  

contin_col_HA_Yes <- filter(contin_col_HA, Heart_Attack_Occurrence == "Yes")

contin_col_HA_Long<-contin_col_HA %>%
  pivot_longer(cols = -c(Heart_Attack_Occurrence, Age_Grp), names_to = "Field", values_to = "Value")
               
contin_col_HA_Long_Yes <- filter(contin_col_HA_Long, Heart_Attack_Occurrence == "Yes")

contin_col_HA_Long_youth <- filter(contin_col_HA_Long, Age_Grp == "Youth")  
```

```{r}
#| fig-width: 10 #to widen the space
#| fig-height: 6 #to lengthen the graph

histograms_HA <- ggplot(contin_col_HA_Long, aes(x = Value, fill = Heart_Attack_Occurrence )) + 
  geom_histogram(binwidth = 1, alpha = 0.8) + 
  theme_minimal() + 
  labs(title = "Histograms of Continuous Variables", y = "Frequency")+
  facet_wrap(~ Field, scales = "free") +
  scale_fill_manual(values = c("Yes" = "red", "No" = "#4682B4"))

histograms_HA

```

::: Callout-note
To visualise the continuous variables, the continuous variables were plotted using the Heart data without re-coding. This is to allow us to be able to view this distribution of the continuous variables.

Based on the histogram, the distribution of of the continuous variables for BMI, Cholesterol Level, Diastolic and Systolic blood pressure, stress levels and heart rate seems to be over normal distribution. The overall trend is similar between across the subjects regardless if they had or had not have a heart attack.
:::

#### 3.2.1 Closer look and the Youth Age Group

```{r}
#| fig-width: 10 #to widen the space
#| fig-height: 6 #to lengthen the graph

histograms_HA_youth <- ggplot(contin_col_HA_Long_youth, aes(x = Value, fill = Heart_Attack_Occurrence )) + 
  geom_histogram(binwidth = 1, alpha = 0.8) + 
  theme_minimal() + 
  labs(title = "Histograms of Continuous Variables", y = "Frequency")+
  facet_wrap(~ Field, scales = "free") +
  scale_fill_manual(values = c("Yes" = "#FF1493", "No" = "#87CEEB"))

histograms_HA_youth
```

::: Callout-note
Comparing the occurrence of heart attack in youth, it follows the same distribution as the over all data. It was noted that similar to the overall data, there resembles the normality in the BMI, Cholesterol, Diastolic and systolic blood pressure, heart rate and stress levels.

This maybe logical as these variables are measure of bodily vitals or measurements, and due to homostasis the self regulating ability of the human body it will be normal to be peak in the center.

Hence to better understand variables, it may be more meaning for categorszed them into established bench marks such as what if the rest heart rate above 100 is considered as elevated.
:::

```{r}

#| fig-width: 24 #to widen the space
#| fig-height: 24 #to lengthen the graph

HA_youth %>%
  ExpNumViz(target = "Heart_Attack_Occurrence",
            nlim=10,
            Page=c(2,4))


```

::: Callout-note
Besides histogram, given that the distribution is resemble normal distribution, boxplot could be useful type of chart to visualise if there are any differences across the variables. It was noted that there maybe slight differences in stress levels between the youth with heart attack verse youth without heart attack.
:::

## 4.0 Confirmatory Data Analysis (CDA)

### 4.1 Chi Square test for youth data

Due to there is no trend observed by examining the barcharts, the chi square was ran using all variables.

After running the above associative test, the following code chunk to carry out the chi-square test to find our variables with association between heart attacks and categorical variables. The following codes on how to perform chi-square is generated using co-pilot.

```{r}
analyze_association <- function(data, x, y) {
  # Perform Chi-Square test
  table <- table(data[[x]], data[[y]])
  chi_test <- chisq.test(table)
  
  # Print Chi-Square test results
  print(paste("Chi-Square Test between", x, "and", y))
  print(chi_test)
}

# Analyze association for each categorical column with Heart_Attack_Occurrence
for (column in colnames(categorical_columns)) {
  if (!(column %in% c("Heart_Attack_Occurrence", "Event_Grp", "Age_Grp")))  
    analyze_association(HA_youth, column, "Heart_Attack_Occurrence")
  }



```

### 4.2 Significant Test of Association (Dependence) : ggbarstats() methods

```{r}


A_HA_youth_BMI <- ggbarstats(HA_youth, 
           x=BMI_cat,
           y=Heart_Attack_Occurrence)

A_HA_youth_Dia <- ggbarstats(HA_youth, 
           x=Diabetes_History,
           y=Heart_Attack_Occurrence)


```

```{r}

#| eval: false
#| fig-width: 120 #to widen the space
#| fig-height: 120 #to lengthen the graph

Patchwork_sig <- (A_HA_youth_Dia | A_HA_youth_BMI)


Patchwork_sig 
```

::: Callout-note
Using the ggbarstats, we are able to view the associations between diabetes history and BMI verse heart attack occurrence in youth.

For **Diabetes history**, it is clear that youth with diabetes history have higher association or may have high risk with heart attack as compared to youth without diabetes in the past.

For the **BMI category**, while the statistical test seems that BMI category may be significant, however it seems like the proportion of youth with normal BMI may result in high proportion of youth has heart attack. While it is consistent with general knowledge, it is noted the youth who are underweight has slightly less association of having heart attack.

Further analysis using logistic regression would be required to further understand the relationship.
:::

### 4.3 Chi Square for subjects with heart attack of the two different age group

```{r}
analyze_association <- function(data, x, y) {
  # Perform Chi-Square test
  table <- table(data[[x]], data[[y]])
  chi_test <- chisq.test(table)
  
  # Print Chi-Square test results
  print(paste("Chi-Square Test between", x, "and", y))
  print(chi_test)
}

# Analyze association for each categorical column with Heart_Attack_Occurrence
for (column in colnames(categorical_columns)) {
  if (!(column %in% c("Heart_Attack_Occurrence", "Event_Grp", "Age_Grp")))  
    analyze_association(HA_Yes, column, "Age_Grp")
}

```

::: Callout-note
Based on chi square test, there no statistically significant variables for the between the categorical variables and age group. Hence the earlier observations in section 3.1.1 are not valid.
:::

### 4.5 Anova performed on the youth group.

Anova is performed to test the differences in true mean across continous variables for the youth with heart attack and without heart attack. The following code is steamlined with the help of Microsoft co-pilot.

```{r}
perform_anova <- function(data, group_var, continuous_vars) {
  results <- list()
  
  for (var in continuous_vars) {
    print(paste("Performing ANOVA for", var, "by", group_var))
    anova_result <- summary(aov(as.formula(paste(var, "~", group_var)), data = data))
    results[[var]] <- anova_result
    print(anova_result)
  }
  
  return(results)
}

# List of continuous variables
continuous_vars <- c("BMI", "Cholesterol_Level", "Stress_Levels", "Heart_Rate", "Systolic_BP", "Diastolic_BP")

# Perform ANOVA for each continuous variable in the Youth group
anova_results <- perform_anova(contin_col_HA_youth, "Heart_Attack_Occurrence", continuous_vars)

```

#### 4.5.1 Oneway ANOVA Test: ggbetweenstats() method

```{r}
stress <- ggbetweenstats(
  data = HA_youth,
  x = Heart_Attack_Occurrence, 
  y = Stress_Levels,
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "s",
  p.adjust.method = "fdr",
  messages = FALSE
)

stress
```

::: callout-note
Based on the anova results, there is statistically significant differences in the stress levels between youth with heart attack verse youth whom have not had heart attack.

However, it is noted that there small effect size (Hedges' g = 0.09) suggests that although the difference is statistically significant, but the small effect may not may not be large enough to have meaningful implications in a real-world context.

As the stress level maybe derived from Likert kind of survey, it maybe meaningful to categorise it to the different levels and analyse using logistic regression model. This had been done in section 2.6
:::

## 5.0 Model building

Due to the lack of observable trends from CDA and EDA, explanatory model is build to further analysis if there are other variables, besides Stress level, BMI categories and Diabetes history, a logistic model is built to further analyse the data set for trends.

### 5.1 Logistic Regression Model

```{r}
# Fit the model
model_log_youth <- glm(Heart_Attack_Occurrence_num  ~ Gender_num + Region_num + Smoking_History_num + Diabetes_History_num + Physical_Activity_num + Diet_Quality_num + Alcohol_Consumption_num + Stress_Level_cat_num + BMI_cat_num + Heart_Rate_cat_num + BP_cat_num +Family_History_num + Cholesterol_Level_cat_num, family = binomial(), data = HA_youth)

# View the summary
summary(model_log_youth)
```

```{r}
exp(coef(model_log_youth))
```

```{r}
#| eval: false
#| fig-width: 24 #to widen the space
#| fig-height: 24 #to lengthen the graph

# Check collinearity
check_c <- check_collinearity(model_log_youth)

# Print the collinearity check results
print(check_c)

# Plot collinearity diagnostics
plot(check_c)
```

```{r}
confint.default(model_log_youth)
```

```{r}
library(caret)
predictions <- predict(model_log_youth, type = "response")
predicted_classes <- ifelse(predictions > 0.5, 1, 0)
confusionMatrix(as.factor(predicted_classes), as.factor(HA_youth$Heart_Attack_Occurrence_num))
```

### 5.2 Variables important

```{r}
# Extract coefficients from the model
coef_data <- as.data.frame(coef(summary(model_log_youth)))
coef_data$Feature <- rownames(coef_data)
rownames(coef_data) <- NULL

# Plot the coefficients
library(ggplot2)
ggplot(coef_data, aes(x = reorder(Feature, Estimate), y = Estimate)) +
  geom_bar(stat = "identity", fill = "#FF1493") +
  coord_flip() +
  labs(title = "Feature Importance (Logistic Regression Coefficients)",
       x = "Feature",
       y = "Coefficient Estimate") +
  theme_minimal()
```

::: Callout-note
The model results show that stress level and diabetes history are statically significant in leading to higher risk of heart attack in youth.

Besides two two variables, it seems that Diet quality may also play a role in leading heart attack in Japanese youth, although the significant level at 0.1 which is more than the standard 0.05.
:::

#### 5.2.1 other codes in looking into model parameters (should not be include in the assignment counts) 

```{r}
library(caret)
predictions <- predict(model_log_youth, type = "response")
predicted_classes <- ifelse(predictions > 0.5, 1, 0)
confusionMatrix(as.factor(predicted_classes), as.factor(HA_youth$Heart_Attack_Occurrence_num))
```

```{r}
set.seed(123)
train_indices <- sample(1:nrow(HA_youth), 0.7 * nrow(HA_youth))
train_data <- HA_youth[train_indices, ]
test_data <- HA_youth[-train_indices, ]

# Fit the model on the training data
model_train <- glm(Heart_Attack_Occurrence_num ~ Gender_num + Region_num + Smoking_History_num + Diabetes_History_num + Physical_Activity_num + Diet_Quality_num + Alcohol_Consumption_num + Stress_Level_cat_num + BMI_cat_num + Heart_Rate_cat_num + BP_cat_num + Family_History_num + Cholesterol_Level_cat_num, family = binomial(), data = train_data)

# Predict on the test data
test_predictions <- predict(model_train, newdata = test_data, type = "response")
test_predicted_classes <- ifelse(test_predictions > 0.5, 1, 0)

# Evaluate the model on the test data
confusionMatrix(as.factor(test_predicted_classes), as.factor(test_data$Heart_Attack_Occurrence_num))

```

```{r}
#| eval: false
ggcoefstats(model_log_youth, 
            output = "plot")
```

```{r}

AIC(model_log_youth )
```

## 6.0 Final charts and conclusion

::: Call-note
With acknowledge of the how the variables related to the occurrence of heart attack, in this section, we will do up the charts to visually show the relationship between the variables and heart attack in youth.

Since the earlier charts has already compared the main group of of interest youth with heart attack, with adults with heart attack and youth without heart attack. This section is dedicated to show visualise the relationship of the variables within the group
:::

```{r}

P1 <-ggplot(data=HA_youth_yes, 
            aes(x= BMI, 
                y= Stress_Levels, 
                color=Diabetes_History)) +
  geom_point(size=1) +
  coord_cartesian(xlim=c(5,40),
                  ylim=c(0,11))

P2<- ggMarginal(P1, type = "boxplot",
           groupColour = TRUE, 
           groupFill = TRUE)+ theme_minimal()
```

```{r}

P3 <-ggplot(data=HA_youth_yes,
       aes(x = Diabetes_History , y = Stress_Levels, color = Diet_Quality )) +
    geom_boxplot()+
  facet_wrap(~ BMI_cat)+ theme_minimal()

```

```{r}
P4<-ggplot(HA_youth_yes, 
       aes(x = Diabetes_History, y = Stress_Levels)) +
  stat_halfeye(adjust = 0.5,
               justification = 0.1,
               .width = 0,
               point_colour = NA) +
  stat_dots(side = "left", 
            justification = 1.1, 
            binwidth = .5,
            dotsize = 0.05)
  #coord_flip() +
  theme_economist()
```

```{r}


#| fig-width: 24 #to widen the space
#| fig-height: 24 #to lengthen the graph

Patchwork <- P4 |P3

Patchwork
```

::: Callout-note
After establishing the following:

-Stress Level and Diabetes History have statistically significant related to incidents of heart attack in youth using the logistic regression model. Although diet history may not have statistically significant, but is worthwhile to explore its relationship with other variables.

-there is statistically significant differences in true group mean in stress level between youth whom had heart attack and without heart attack.

-the various BMI categories seem to have statically significant associations in relation of heart attack, proven using chi square.

The 4 variables of the ***youth whom had heart attack*** plotted using the half eye chart, box plot across the BMI categories to further analyse the relationships among the mentioned variables.

**Observation**

As mentioned in section 5, that the logistic regression, that the stress level is by Likert scale as the data is discrete instead of continuous as shown in the dot stats chart. Based on the chart, we can see the youth with higher stress level had diabetes before as compared to those who has not have diabetes before.

This trend could be further showed by the box plot on the right. The boxplot is facet across the various BMI categories.

At a quick glance, we could see obese youth whom had heart attack, has highest stress level. This particular group of youth also had the poor diet quality and diabetes before.

It was also noted there are overall higher stress level in overweight and obese youths as compared to those with normal BMI or underweight.

Among the youths, it was also noted. that two groups with the highest stress level are the obese and overweight youths with poor diet quality.

Further analysis could be carried out to further explore characteristic of Japanese Youth whom had heart attack. It is important to note that more information on the data set would be require before further conclusion could be drawn

**Conclusion**

Learning points from this take home exercise is that data visualising is not direct. It requires student or analyst to have a throughout understanding of the data set before the analyst could make use of the tools to present the data in a accurate and simplify manner which allow viewers to be able to see the trend of the data set.
:::

### 7.0 Other Charts and visualisation tools used.

As part of the learning journey, there are may other charts tried to visualise the data. However due to the assignment limit, they are not added to the main part of the assessment.

##### Histogram to compare subjects with heart attack in youth and adults

```{R}

histograms <- ggplot(contin_col_HA_Long_Yes, aes(x = Value, fill = Age_Grp )) + 
  geom_histogram(binwidth = 1, alpha = 0.8) + 
  theme_minimal() + 
  labs(title = "Histograms of Continuous Fields", y = "Frequency")+
  facet_wrap(~ Field, scales = "free") 

# Print the histograms
print(histograms)
```

##### visualizing using density plot to further understand stress level

```{r}
# Adding density plot
ggplot(HA_youth, aes(x = Stress_Levels)) +
  geom_histogram(aes(y = ..density..), binwidth = 1, fill = "blue", alpha = 0.5) +
  geom_density(color = "red", size = 1) +
  labs(title = "Histogram and Density Plot of Stress Levels", x = "Stress Levels", y = "Density")
```

##### Correlation plots

```{r}
HA.cor <- cor(HA_youth_yes[, c(7, 11:15)], use = "complete.obs")
```

```{r}
corrplot(HA.cor)
```

```{r}

ggplot(HA, aes(x = Stress_Levels, fill = Heart_Attack_Occurrence)) +  # Map fill to Heart_Attack_Occurrence
  geom_histogram(aes(y = ..density..), binwidth = 1, alpha = 0.5, position = "identity") +  # Adjust position for overlap
  geom_density(color = "red", size = 1, alpha = 0.3) +  # Density plot with transparency
  labs(title = "Histogram and Density Plot of Stress Levels",
       x = "Stress Levels",
       y = "Density") +
  theme_minimal(base_size = 14) +  # Improve overall theme and font size
  scale_fill_brewer(palette = "Set2", name = "Heart Attack Occurrence") +  # Use a color palette for filling
  theme(legend.position = "top")  # Position the legend at the top

```

```{r}
scatter_plot <- ggplot(HA, aes(x = Stress_Levels, y = BMI, color = Heart_Attack_Occurrence)) +
  geom_point(size = 1, alpha = 0.4) +  # Adjust point size and transparency
  labs(title = "Scatter Plot of Stress Levels vs. BMI",
       x = "Stress Levels",
       y = "BMI",
       color = "Heart Attack Occurrence") +  # Add title and labels
  theme_minimal(base_size = 14) +  # Use a clean theme with larger base font size
  scale_color_manual(values = c("blue", "red"))  # Customize colors if needed

# Print the scatter plot
print(scatter_plot)

```

```{r}
ggplot(HA, aes(x = Smoking_History, fill = Heart_Attack_Occurrence)) +
  geom_bar(position = "fill") +
  labs(title = "Smoking History vs Heart Attack Occurrence", 
       y = "Proportion", 
       x = "Smoking History") +
  scale_fill_manual(values = c("lightblue", "salmon"), name = "Heart Attack Occurrence")
```

#### Disclaimer

This take home exercise was done with reference to methods in Prof Kam's e text book https://r4va.netlify.app/ , materials on R available online and the use of Microsoft co-pilot to troubleshoot syntax errors and streamline codes.
